---
title: "STA 9701 Project 1"
author: "Lily (Lizheng) Zhou"
date: "October 28, 2019"
output:
  pdf_document: 
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Many of us may like to use `auto.arima()` function to fit best ARIMA model for univariate time series. This function returns a decent ARIMA model according to either AIC, AICc or BIC value. It is fast, easy and convenient to use, can provide us candidate models or give us suggestions of p and q values according to smallest AIC, AICc or BIC values. It saved much of our time. 

However, I wonder if this function always provides well models. Imagine such a situation, if an original dataset has both seasonal pattern and other patterns, while seasonal pattern is not that strong, I am curious that whether `auto.arima()` would choose to keep seasonal pattern or not. 

To further investigate this question, I need a seasonal dataset but this seasonality should not be that strong. 

There is a great time series dataset package called "Time Series Data Library (tsdl)" (<https://pkg.yangzhuoranyang.com/tsdl/>), which contains a list of 648 time series datasets. I used a `for` loop to auto print the datasets, dataset attributes and plot the datasets. Finally, I found that the 97th dataset in the package fulfills my need. 

```{r preparation, include=FALSE}
# Load library
library(forecast)
library(astsa)
library(tseries)
library(tsdl)
```

# Exploratory Data Analysis 

The 97th dataset of **tsdl** package is: Basic quarterly iron production in Australia: thousand tonnes from March 1956 to September 1994, giving a total of n = 155 observations. Source: Australian Bureau of Statistics (<https://www.abs.gov.au/>)

```{r datasetplot, fig.height=6.06, fig.width=10, echo=FALSE}
# Dataset plot
par(mfrow=c(1,1))
x = tsdl[[97]]
autoplot(x,main='Quarterly Iron Production in Australia: Mar 1956 - Sep 1994')
```

When look at the plot of original data plot, we see that there is a seasonal pattern, wile this seasonality is not that strong, this satisfies my need. 

The time series plot indicates that there are some the variance differences, so original may not be stationary. To investigate the required transformation, we use BoxCox procedure. `BoxCox.lambda()` function gives us lambda = `r BoxCox.lambda(x)` which is close to -0.5. Thus, we take 1 over square root transformation. 

```{r Transformation, results=FALSE, echo=FALSE, fig.height=5, fig.width=10}
# Transformation
BoxCox.lambda(x)
tx = 1/(sqrt(x))
plot(tx,main='Transformed Original Data') 
```

Also took a first order diferencing according to the trend of original data. 

```{r Differencing, echo=FALSE, fig.height=5, fig.width=10}
# Differencing
dtx = diff(tx) 
plot(dtx,main='Differenced of Transformed Original Data')
```

To test if the differenced of transformed original data is stationary, use `adf.test()` function. 

```{r adf test, echo=FALSE, message=FALSE, warning=FALSE}
# adf test
adf.test(dtx) 
```

From the result, we see it is stationary. Conbine with the plot shown above, I decide to use differenced of transformed original data (dtx) to do further analysis. 

The sample ACF and sample PACF of differenced of the transformed data are
shown below: 

```{r ACF and PACF of dtx, echo=FALSE, fig.height=8, fig.width=10}
# ACF and PACF of dtx
par(mfrow=c(2,1))
acf(dtx,155,main='ACF of differenced of transformed data')
pacf(dtx,155,main='PACF of differenced of transformed data')
```

There is a seasonal pattern, while most are not significant, so I feel ACF and PACF all cut off actually. Thus, the ACF and PACF plots suggest a tentative model may be an AR(3), AR(4), MA(2), MA(4) of differenced of transformed data. 

# ARIMA Modeling
Fit 4 models: AR(3), AR(4), MA(2), MA(4) of differenced of transformed data, AIC and BIC shown below. 

```{r fit the model and try AIC and BIC, include=FALSE}
# ARIMA test
ARIMA1 = arima(dtx, order=c(3,0,0))
ARIMA2 = arima(dtx, order=c(4,0,0))
ARIMA3 = arima(dtx, order=c(0,0,2))
ARIMA4 = arima(dtx, order=c(0,0,4))
# Check model AIC and BIC
AIC(ARIMA1);BIC(ARIMA1)
AIC(ARIMA2);BIC(ARIMA2)
AIC(ARIMA3);BIC(ARIMA3)
AIC(ARIMA4);BIC(ARIMA4)
```

::: l-body-outset
|Model|AIC                     |BIC                     |
|-----|------------------------|------------------------|
|AR(3)|`r round(AIC(ARIMA1),3)`|`r round(BIC(ARIMA1),3)`| 
|AR(4)|`r round(AIC(ARIMA2),3)`|`r round(BIC(ARIMA2),3)`| 
|MA(2)|`r round(AIC(ARIMA3),3)`|`r round(BIC(ARIMA3),3)`| 
|MA(4)|`r round(AIC(ARIMA4),3)`|`r round(BIC(ARIMA4),3)`| 
:::

we notice that AIC and BIC both are very close, respectively. I simply choose the smallest BIC here, so the model is AR(3) of differenced of transformed data. 

At first, fit the AR(3) model of differenced of transformed data with constant, the estimated coeffcients and the corresponding p-values are given by `sarima()` function: 

```{r AR3 with constant, include=FALSE}
# AR3 with constant 
summary(ARIMA1)
sarima(dtx, 3,0,0)
```

::: l-body-outset
| Parameter    | Estimate| Standard Error |t-value| p-value|
|--------------|---------|--------|---------|--------|
|$\hat{\phi_1}$| -0.3619 | 0.0780 | -4.6428 | 0.0000 |
|$\hat{\phi_2}$| -0.3181 | 0.0788 | -4.0367 | 0.0001 |
|$\hat{\phi_3}$| -0.2470 | 0.0777 | -3.1805 | 0.0018 |
|$\hat{\mu}$   | -0.0001 | 0.0001 | -1.4597 | 0.1465 |
:::

Since the constant term is not significant, re-fit the model without constant. The following Table shows
the parameter estimates of the model of AR(3) of differenced of transformed data without constant.

```{r AR3 without constant, include=FALSE}
# AR3 without constant 
AR3_no_c = arima(dtx, order=c(3,0,0), include.mean=FALSE)
summary(AR3_no_c)
sarima(dtx, 3,0,0, no.constant=TRUE) 
AIC(AR3_no_c);BIC(AR3_no_c)
```

::: l-body-outset
| Parameter    | Estimate| Standard Error |t-value| p-value|
|--------------|---------|--------|---------|--------|
|$\hat{\phi_1}$| -0.3240 | 0.0787 | -4.1148 | 0.0001 |
|$\hat{\phi_2}$| -0.2789 | 0.0796 | -3.5061 | 0.0006 |
|$\hat{\phi_3}$| -0.2105 | 0.0785 | -2.6796 | 0.0082 |
:::

# `auto.arima()` Modeling
Go back to the purpose of this project, I want to know if `auto.arima` function would always keep the seasonal pattern when seasonal pattern is not strong. Thus, use `auto.arima` function to get the model. 

```{r auto.arima function, echo = FALSE} 
# Use auto.arima function
ARIMAfit_dtx = auto.arima(dtx, approximation=FALSE,trace=FALSE)
summary(ARIMAfit_dtx) 
```

We see that although we already know that this dataset doesn't have strong seasonality pattern, but `auto.arima` function still chooses to provide us a seasonal model ARIMA(0,1,2)(1,0,0)[4]. 

I will compare this seasonal model **ARIMA(0,1,2)(1,0,0)[4]** with the model **ARIMA(3,1,0) without constant** we fit manually. 

# Model Comparison
## AIC and BIC Comparison 
we first compare the AIC and BIC of two models. 

```{r AIC BIC compare, include=FALSE}
# AR3 without constant 
AIC(AR3_no_c);BIC(AR3_no_c) 
# ARIMA(0,1,2)(1,0,0)[4] 
AIC(ARIMAfit_dtx);BIC(ARIMAfit_dtx)
```

::: l-body-outset
| Model                         | AIC                            | BIC                            |
|-------------------------------|--------------------------------|--------------------------------|
| ARIMA(3,1,0) without constant | `r round(AIC(AR3_no_c),3)`     | `r round(BIC(AR3_no_c),3)`     |
| ARIMA(0,1,2)(1,0,0)[4]        | `r round(AIC(ARIMAfit_dtx),3)` | `r round(BIC(ARIMAfit_dtx),3)` | 
:::

I feel that two models' AIC and BIC are still very close, respectively. 

## Diagnostic Checking and Comparing 
Then, compare the diagnostic plots two models. Diagnostic plots shown below: 

```{r Compare Diagnostic plots, message=FALSE, warning=FALSE, echo=FALSE, results=FALSE} 
# Compare Diagnostic plots
sarima(dtx, 3,0,0, no.constant=TRUE)
sarima(dtx, 0,0,2, 1,0,0, 4)
```

Look at the two diagnostic plots, we see that: **Standardized Residuals** totally same; **ACF of Residuals** are all not significant far away from 0; **Normal Q-Q Plot of Std Residuals** look similar, most parts are ok while all have outliers because of 1980s; **p values for Ljung-Box statistic** have no problems, all significant. 

Thus, two diagnostic plots of two models also are very similar and satisfying. 

## Forecasting Comparison 
I use the two models to forecast the differenced of transformed data for next 4 quarters.

**ARIMA(3,1,0) without constant** forecasting plot: 

```{r Forecasting Comparison AR3, message=FALSE, warning=FALSE, echo=FALSE, results=FALSE, fig.height=3.7, fig.width=10} 
# AR(3) forecasting
sarima.for(dtx, 4, 3,0,0, no.constant=TRUE)
```

**ARIMA(0,1,2)(1,0,0)[4]** forecasting plot: 

```{r Forecasting Comparison Seasonal model, message=FALSE, warning=FALSE, echo=FALSE, results=FALSE, fig.height=3.7, fig.width=10} 
# ARIMA(0,1,2)(1,0,0)[4] forecasting
sarima.for(dtx, 4, 0,0,2, 1,0,0, 4)
```

Compare with forecasting of two models, I found that they are still very close, magnitudes are both very small. 

I compared the AIC, BIC, diagnostic plots, forecasting of two models, I feel it is not necessary to keep the seasonal pattern for this dataset, the model `auto.arima()` function provides is kind of complex, AR model is enough. 

# Conclusion 
From this proect, I used a seasonal dataset while seasonality is not strong to do a standard arima model building and get an AR model, later compare with the seasonal model got from `auto.arima()` function. By comparing with the AIC, BIC, diagnostic plots and forecasting of two models, I feel that we don't have to keep seasona pattern for this mode, thus I feel `auto.arima()` function might not always provide the most appropriate and simplest model. 

# Appendix 
## Dataset 

```{r dataset, echo=FALSE}
# Dataset
x
```

## Dataset Attribute 

```{r Attribute, echo=FALSE}
# Dataset Attribute
attributes(x) 
```

## R Code

```{r ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE}
```
